---
title: "Tutorial"
subtitle: "Using R to analyze international large-scale educational assessment data"
author: "HJ WU & CY Chang"
institute: "Public Health & Education"
date: "2020/06/01"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
# Outline

## Why we chose this topic (Chang)    
## Import data & Data wrangling (Wu)
#### Select and merge data 
#### Data wrangling 
## Descriptive analysis(Chang)  
#### Mean and Frequency  
## Regression analysis  (Wu)
## Data visualization  (Wu)

---
class: inverse, center

# Why we chose this topic

## What is PISA?

### 1. The international evaluation execute by OECD (經濟合作暨發展組織)

### 2. Object: 15 years Students. (math, read, science)
--

## Why we chose this topic?

### 1. Researchers often analyze PISA to understand students' learning environment. Take this result to improve education.

### 2. Practice how to analyze Big data.

---

# Import data & Data wrangling (1)

### Select data and merge 
  
- **Install the** [**EdSurvey**](https://www.rdocumentation.org/packages/EdSurvey/versions/2.4.0) **package:**

```{r eval=FALSE, tidy=FALSE}
install.packages("EdSurvey")

```
--

**download, unzip and read the PISA data**
```{r message = FALSE, warning=FALSE}
library(EdSurvey)
downloadPISA(years = 2012, database ="INT", 
             root="C:/Users/X510/Desktop/PISA-2012") 

pisa12<-readPISA(path = "C:/Users/X510/Desktop/PISA-2012/PISA/2012", 
                 database = "INT", 
                 countries = c("AUS","CAN","DEU","ESP","FIN","FRA","GBR",
                               "IDN","ITA","JPN","MEX","MYS","NZL","TAP","USA"), 
                 verbose = T)
```

---
# Import data & Data wrangling (2)  
### Select data and merge 

**Select variable**
```{r}
pisa12raw<-getData(pisa12, 
                   varnames=c("cnt","schoolid","stidstd","questid","bookid",
                              "st04q01","st28q01","st13q01","st26q01","st26q02",
                              "st26q03","st26q04","st26q05","st26q06","st26q07",
                              "st26q08","st26q09","st26q10","st26q11","st26q12",
                              "st26q13","st26q14","w_fstuwt","math", "macc", 
                              "macq", "macs", "macu", "mape", "mapf", "mapi",
                              "read", "scie"), 
               dropUnusedLevels = TRUE, # keep unusedLevels
               omittedLevels = FALSE,   # keep Levels
               defaultConditions = TRUE,
               formula = NULL,
               recode = NULL,
               includeNaLabel = FALSE,
               addAttributes = FALSE)
```

```{r echo=FALSE, comment='#'}

```

---

# Import data & Data wrangling (3)  
## Data wrangling

**Turn to data frame**
```{r comment='#', message=FALSE, warning=FALSE}
pisa.dt<-do.call(rbind, pisa12raw)
str(pisa.dt, list.len=8)
```
---
# Import data & Data wrangling (4)  
## Data wrangling

**Rename the variable**
```{r comment='#', message=FALSE, warning=FALSE}
# aligned with the format of intsvy package
names(pisa.dt) <- toupper(names(pisa.dt))
# dealing with different NA value
levels(pisa.dt$ST28Q01)[levels(pisa.dt$ST28Q01)%in% 
                          c("N/A", "Invalid", "Missing")]<-NA
levels(pisa.dt$ST13Q01)[levels(pisa.dt$ST13Q01)%in% 
                          c("N/A", "Invalid", "Missing")]<-NA
```

---
# Import data & Data wrangling (5)  
## Data wrangling
**Recode**
```{r comment='#', message=FALSE, warning=FALSE}
library(dplyr)
pisa.dt[, 6:22]<-sapply(pisa.dt[ ,6:22], function(x) as.numeric(x)) 
pisa.dt<-pisa.dt%>%mutate(Country_ID=as.factor(.$CNT),
         Gender=factor(abs(as.numeric(.$ST04Q01)-1), 
                      levels=c(0,1), labels=c("F","M")),
         Book=as.numeric(.$ST28Q01),
         Mother=abs(.$ST13Q01-6))

```

--

**orginal code**
```{r eval=FALSE, message=FALSE, warning=FALSE, comment='#'}

dta$CNT<-as.factor(dta$CNT)
for(i in 4:20) {
dta[,i] <- as.numeric(dta[,i])
} 
dta$Gender<-dta$ST04Q01
dta$Gender<-factor(dta$Gender, levels=c(0,1), labels=c("F","M"))
dta$Book<-dta$ST28Q01 
dta$Mother<-dta$ST13Q01

```

---
# Import data & Data wrangling (6)  
## Data wrangling
**Create new variables:posseission **

**Revised code**
```{r comment='#', message=FALSE, warning=FALSE}
# stq2601-stq2614  recode "Yes"==1, "No"==0 
cols.num <- c(paste0(rep("ST26Q0", 9),seq(1,9,by=1), sep=""),
              paste0(rep("ST26Q", 5),seq(10,14,by=1), sep=""))
pisa.dt[cols.num] <-sapply(pisa.dt[cols.num], function(x) abs(x-2))
# Possessions is sum of ST26Q01 to ST26Q14
pisa.dt$Possessions<-rowSums(pisa.dt[cols.num])
```
--
**Original code**
```{r comment='#', message=FALSE, warning=FALSE, eval=FALSE}
for(i in 6:19) {dta[,i] <- abs(dta[,i]-2)}
dta$Possessions<-dta$ST26Q01+dta$ST26Q02+dta$ST26Q03+dta$ST26Q04
+dta$ST26Q05+dta$ST26Q06+dta$ST26Q07+dta$ST26Q08+dta$ST26Q09
+dta$ST26Q10+dta$ST26Q11+dta$ST26Q12+dta$ST26Q13+dta$ST26Q14 
```

---

# Descriptive analysis 
## Frequency
**install package "intsvy", it can produce descriptive table**
```{r, message=FALSE}
# install the package, it can manipulate PISA data
library(intsvy)
```

--

**pisa.table: Produce frequency table according to variables**
```{r}
# Produce frequency table according to Country & Gender
ptableCB <- pisa.table(variable="Book", by=c("Country_ID", "Gender"), data=pisa.dt)
head(ptableCB)# Show data
```

---

```{r}
# alternative table display
knitr::kable(head(ptableCB), align = "cc", format = "html") # align to center
```

---

# Descriptive analysis 
## Frequency
**plot: Plot the frequency table**

.pull-left[
```{r plot-first, fig.show = 'hide'}
plot(na.omit(ptableCB), stacked=T)
#stacked=T: Stack the histogram
#na.omit: Omit NA
```
]

.pull-right[
```{r ref.label = 'plot-first', echo = FALSE}
```
]

--

德國(DEU)：The most students whose Book collection > 500. On the countrary: 墨西哥(MEX).

---

# Descriptive analysis 
## Mean
**pisa.mean.pv: Produce mean table according to variables**

```{r echo=TRUE, cache=TRUE}

# Produce mean table according to Country & Gender & Book
pmeansMCGB <- pisa.mean.pv(pvlabel="MATH",
                           by=c("Country_ID", "Gender", "Book"),
                           data=pisa.dt,
                           export=FALSE)# output csv. file
head(ptableCB)#show data
```

---

# Descriptive analysis 
## Mean
**plot: Plot the mean table**

.pull-left[
```{r plot-second, fig.show = 'hide'}
plot(na.omit(pmeansMCGB), sort=TRUE)
#sort: list data in order
#na.omit: Omit NA
```
]

.pull-right[
```{r ref.label = 'plot-second', echo = FALSE}
```
]

--

Tanwan's math score are higher than other countries.
On the countrary: Indonesia.

---

# Descriptive analysis 
## Descriptive table
**install package "arsenal", it can produce descriptive table**
```{r echo=TRUE, cache=TRUE}
library(arsenal)
library(DT)
```

```{r echo=TRUE, cache=TRUE}
# Choose variables which you want to put in the table.
tab_select <- pisa.dt[,c("Country_ID","Gender","Book","Possessions","Mother")]
str(tab_select)
```

---

# Descriptive analysis 
## Descriptive table

**Setting**
```{r echo=TRUE, cache=TRUE}
controls <- tableby.control(
  total = T,
  numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),#setting numeric variables 
  cat.stats = c("countpct", "Nmiss2"),#setting category variables 
  stats.labels = list(
   meansd = "Mean (SD)",
   medianq1q3 = "Median (Q1, Q3)",
   range = "Min - Max",
   Nmiss2 = "Missing"))
```

---

# Descriptive analysis 
## Descriptive table
**Show table**
```{r echo=TRUE, cache=TRUE}
table <- tableby(Country_ID ~ ., data = tab_select, control = controls) #Categorize by Country

summary(table, text=TRUE, title = "Descriptive table by Country")
```

---
class: inverse, center, middle

# Regression analysis  (Wu)

---
# Regression analysis (1)
**regression model** 
```{r message=FALSE,warning=FALSE, fig.retina=2, out.width=5 }
library(intsvy)
rmodelMGBMP <-pisa.reg.pv(pvlabel="MATH",
                          x=c("Gender","Book","Mother","Possessions"), 
                          by = "Country_ID", data=pisa.dt,
                          export=FALSE)

```
---

# Regression analysis (2)
```{r}
plot(rmodelMGBMP)
```

---
# Regression analysis (3)
**Extract estimate from regression model for further data visulization**
### 1
```{r}
model.e<-sapply(rmodelMGBMP, "[", 5)
model.e1<-bind_rows(model.e)%>%mutate(country=rep(c("AUS","CAN","DEU","ESP","FIN","FRA","GBR","IDN",
                                              "ITA","JPN","MEX","MYS","NZL","TWN","USA"), each=6),
                                item=rep(rownames(model.e$Australia.reg), 15))
knitr::kable(head(model.e1), format = "html")
```

---
**Extract estimate from regression model for further data visulization**

```{r}
# alternative method
datalist = list()
for (i in 1:15){datalist[[i]]<-rmodelMGBMP[[i]]$reg}
 model<-do.call(rbind, datalist)%>%
  mutate(Country=rep(c("AUS","CAN","DEU","ESP","FIN","FRA","GBR","IDN","ITA","JPN","MEX","MYS", "NZL","TWN","USA"), each=6)) 
model$variable<-rep(rownames(rmodelMGBMP[[1]]$reg), 15)
knitr::kable(head(model), format = "html")
```

---
**Extract estimate from regression model for further data visulization**
```{r warning=FALSE, ,message=FALSE}
library(dplyr)
model.s<-model.e1%>%
  dplyr::select(item, country, Estimate)%>%
  tidyr::spread(item, Estimate)
model.s[order(model.s$GenderM),]
```

---
```{r warning=FALSE, ,message=FALSE}
model.s$genderlevel<-with(model.s, 
                          factor(findInterval(GenderM, c(-Inf,quantile(GenderM, probs=c(0.25, 0.75)),Inf)), labels=c("後面25%", "中間50%", "前面25%"))) 

knitr::kable(head(model.s), format = "html")
```
---
class: inverse, center
# Data visulization 

---
# Data visulization
```{r ,message=FALSE, warning=FALSE}
library(rworldmap)
library(countrycode)
library(RColorBrewer)
```

**mapCountryData**
```{r ,message=FALSE, warning=FALSE}
cols<-heat.colors(6)
genderMap <- joinCountryData2Map(model.s, joinCode = "ISO3", nameJoinColumn = "country")

```

---

# Data visulization
**mapCountryData**
.pull-left[
```{r plot-third, fig.show = 'hide', message=F, warning=F}

mapCountryData(genderMap, nameColumnToPlot = "genderlevel", catMethod = "",
               addLegend = TRUE, missingCountryCol = "white",
               colourPalette = rev(heat.colors(6)),
               mapTitle = 'Gender estimate level in PISA 2012 among 15 countries')

```
]

.pull-right[
```{r ref.label = 'plot-third', echo = FALSE, message=F, warning=F}
```
]

---
# Data visulization
**ggplot**
```{r ,message=FALSE, warning=FALSE}
# Data visualization

library(ggplot2)
library(dplyr)
library(mapproj)
# rename countries to meet the map_data
model.sc<-model.s%>%
  mutate(countries=c("Australia", "Canada", "Germany", 
                     "Spain", "Finland", "France", "United Kingdom",
                     "Indonesia", "Italy","Japan", "Mexico", "Malaysia", 
                     "New Zealand", "Taiwan", "USA"))

WorldData <- map_data('world') %>% filter(region != "Antarctica") %>% fortify


```
---
# Data visulization
**ggplot**

.pull-left[
```{r plot-forth, fig.show = 'hide', message=FALSE, warning=FALSE}
ggplot() +
  geom_map(data = WorldData, map = WorldData,
           aes(x = long, y = lat, group = group, map_id=region),
           fill = "white", colour = "#7f7f7f", size=0.5) + 
  geom_map(data = model.sc, map=WorldData,
           aes(fill=genderlevel, map_id=countries),
           colour="#7f7f7f", size=0.5) +
  coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90)) +
  scale_fill_manual(values=brewer.pal(6, "Reds")) +

  labs(fill="legend", title="Title", x="", y="") +
  theme_bw()
```
]

.pull-right[
```{r ref.label = 'plot-forth', echo = FALSE, message=F, warning=F}
```
]

---
class: inverse, center

# Conclusion

### We try to simplify r-code, introduce different packages to manipulate the data. There are three useful packages we use in the task:

### 1. EdSurvey：It can read in and analyze education surveys. e.g.PISA, TIMSS.

### 2. intsvy：It can manipulate PISA data.

### 3. arsenal：Use for large-scale statistical summaries.

---
class: inverse, center, middle

# Thank you for listening
